---
output: github_document
---

```{r, echo = FALSE}
library(knitr)
```
### Seed banks in Clarkia xantiana

Siegmund, G.-F., and M. A. Geber. 

---

### Abstract

....

---

### Repository Directory

1. `library` contains the [...] datasets associated with the paper. I have included the metadata for each dataset below. 

2. `reshapeDataScripts` contains the R scripts used to download and reshape the datasets.

3. `modelBuild` contains the R scripts to fit models.

4. `modelCodeJAGS` contains the JAGS scripts.

5. `analysisCode` contains scripts to analyze model output.

6. `products` contains a folder with the R scripts used to make all figures and tables, and a folder with the manuscript.

----

### Library Metadata

#### sigmaDF

Metadata for the seedling survival to fruiting dataset described in the Methods and Appendices XXXX.

```{r, echo = FALSE}
load(file = "~/Dropbox/projects/clarkiaScripts/data/cleanData/sigmaDF.RData")
df <- sigmaDF[,c(1:6)]
dt <- data.frame(field = names(df))
dt$description[1] <- "name of site, see methods and Appendix XX"
dt$description[2] <- "year of ..."
dt$description[3] <- "transect identifier, see Methods and Appendix XX"
dt$description[4] <- "position within transect identifier, see Methods and Appendix XX"
dt$description[5] <- "number of seedlings at plot"
dt$description[6] <- "number of fruiting plants at plot"
kable(dt)
```

<br>