---
output: github_document
---

```{r, echo = FALSE}
library(knitr)
```
### Seed banks in Clarkia xantiana

Siegmund, G.-F., and M. A. Geber. 

---

### Abstract

....

---

### Repository Directory

1. `library` contains the [...] datasets associated with the paper. I have included the metadata for each dataset below. 

2. `reshapeDataScripts` contains the R scripts used to download and reshape the datasets.

3. `modelBuild` contains the R scripts to fit models.

4. `modelCodeJAGS` contains the JAGS scripts.

5. `analysisCode` contains scripts to analyze model output.

6. `products` contains a folder with the R scripts used to make all figures and tables, and a folder with the manuscript.

----

### Library Metadata

The table below summarizes the data generated in this study (also available [as a pdf at this link](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/products/tables/data-summary.pdf))

  ---------------------------------------- ------------------ ------------------------ -----------
  **Seed vital rates**                      ---                          ---                ---
  Seed survival and germination            Seed bag burial      Y<sub>1</sub>            2006-2009
  Seed viability                           Viability trials    Y<sub>2</sub>             2006-2009
  Seed survival and germination            Seed pots            Y<sub>3</sub>            2013-2019
  **Seedling survival**                   ---                          ---                ---
  Seedling survival to fruiting            Field surveys        Y<sub>4</sub>            2006-2019
  **Fruits per plant**                       ---                          ---                ---
  Total fruit equivalents per plant        Field surveys       Y<sub>5</sub>             2006-2012
  Undamaged and damaged fruits per plant   Field surveys        Y<sub>6</sub>            2013-2019
  Total fruit equivalents per plant        Extra plots          Y<sub>7</sub>            2006-2012
  Undamaged and damaged fruits per plant   Extra plots          Y<sub>8</sub>            2013-2019
  **Seeds per fruit**                       ---                          ---                ---
  Seeds per undamaged fruit                Lab counts           Y<sub>9</sub>            2006-2019
  Seeds per damaged fruit                  Lab counts          Y<sub>0</sub>             2013-2019
  ---------------------------------------- ------------------ ------------------------ -----------

Links to  all the datasets follow:

* [Seed bag burial](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/library/dataProcessingWorkflow.md#seed-bag-data)
* [Viability trials](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/library/dataProcessingWorkflow.md#viability-trial-data)
* [Seed pots](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/library/dataProcessingWorkflow.md#seed-pot-data) - needs to be updated
* [Field surveys for seedling survival to fruiting](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/library/dataProcessingWorkflow.md#seedlings-and-fruiting-plant-data)
* [Field surveys for total fruit equivalents per plant, and undamaged/damaged fruits per plant](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/library/dataProcessingWorkflow.md#fruits-per-plant-data-extra-plots)
* [Lab counts for seeds per fruit from undamaged and damaged fruits](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/library/dataProcessingWorkflow.md#seeds-per-fruit-data)

### Summary

The table below summarizes the parameters and years for which I have estimates. 

```{r, echo=FALSE}
variables = c("Parameter",2006:2019)

s0 = c("s0", rep("N",3),rep("-",11))
s1 = c("s1", rep("Y",3),rep("-",11))
g1 = c("g1", rep("Y",3),rep("-",11))
s2 = c("s2", rep("Y",3),rep("-",11))
sigma = c("sigma", rep("Y",14),rep("N",0))
Fec = c("Fec", rep("Y",8),rep("N",6))
phi = c("phi", rep("Y",13),rep("N",1))

dt = rbind(s0,s1,g1,s2,sigma,Fec,phi)

dt = data.frame(dt,row.names=dt[,1])

colnames(dt) = variables
kable(dt,row.names=FALSE)
```

#### sigmaDF

Metadata for the seedling survival to fruiting dataset described in the Methods and Appendices XXXX.

```{r, echo = FALSE}
load(file = "~/Dropbox/projects/clarkiaScripts/data/cleanData/sigmaDF.RData")
df <- sigmaDF[,c(1:6)]
dt <- data.frame(field = names(df))
dt$description[1] <- "name of site, see methods and Appendix XX"
dt$description[2] <- "year of ..."
dt$description[3] <- "transect identifier, see Methods and Appendix XX"
dt$description[4] <- "position within transect identifier, see Methods and Appendix XX"
dt$description[5] <- "number of seedlings at plot"
dt$description[6] <- "number of fruiting plants at plot"
kable(dt)
```

<br>