---
output: github_document
---

### Data processing workflow

I'm trying to document the data processing workflow. I am using the packages `tidyverse` and `readxl` (documentation: https://readxl.tidyverse.org/). Link to the Issue documentation for the data processing workflow: https://github.com/gregor-fausto/clarkiaSeedBanks/issues/6. 

Load the libraries for data processing (see https://github.com/r-lib/rlang/issues/669 for the overwrite message I am suppressing)

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(knitr)
library(lubridate)
```

Start with the survival data. Scripts for the original processing I did is located in the following folder. I want to document when I last updated these files before I start updating the workflow. None of the original scripts are going to be changed. 

```{r}
dir = c("/Users/Gregor/Dropbox/projects/clarkiaScripts/code/reshapeDataScripts/")
files = list.files(path=dir)
dt <- data.frame(files)
fileModified <- function(x){file.info(paste0(dir,x))$mtime}

for(i in 1:dim(dt)[1]){
  dt[i,2] <- ymd_hms(fileModified(files[i])) %>%
    data.frame() %>% rename(dateTime='.') 
}

kable(dt)
```

Start with the seed data. This is file `.../reshapeSeeds.R` file in the list above

```{r}
setwd("/Users/Gregor/Dropbox/Clarkia-LTREB/20_demography_sites")

filenames<-list.files(pattern=paste("^seeds_"), recursive=TRUE)
worksheet_2006.2010 <- filenames[1:5]
worksheet_2011.2012 <- filenames[6:7]
worksheet_2013.2015 <- filenames[8:10]
print(worksheet_2006.2010,worksheet_2011.2012,worksheet_2013.2015)
```


