---
output: github_document
---

### Data processing workflow

I'm trying to document the data processing workflow. I am using the packages `tidyverse` and `readxl` (documentation: https://readxl.tidyverse.org/). Link to the Issue documentation for the data processing workflow: https://github.com/gregor-fausto/clarkiaSeedBanks/issues/6. 

- [File directories](#file-directories)
- [Processing data](#processing-data)
- [Seedlings and fruiting plant data](#seedlings-and-fruiting-plant-data)
- [Fruits per plant data for transects](#fruits-per-plant-data-for-transects)
- [Fruits per plant data extra plots](#fruits-per-plant-data-extra-plots)
- [Seeds per fruit data](#seeds-per-fruit-data)
- [Seed bag data](#seed-bag-data)
- [Viability trial data](#viability-trial-data)
- [Seed pot data](#seed-pot-data)

### File directories

The files holding datasets are originally found in the shared Dropbox folder `.../Dropbox/Clarkia-LTREB/20_demography_sites/`. I created a folder on my local Dropbox `/Users/Gregor/Dropbox/dataLibrary/` to hold a copy of these files. To copy the files, I run the following in my Terminal:

`cp /Users/Gregor/Dropbox/Clarkia-LTREB/20_demography_sites/* /Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/`

I also ran: 

`cp /Users/Gregor/Dropbox/Clarkia-LTREB/20_demography_sites/new\ fruit\ \&\ seed\ files/* /Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/new\ fruit\ \&\ seed\ files/`

I last ran these on 04/15/20 at 5:20 PM. The files in this directory are now a copy of those in the shared Dropbox folder. For the time being, the copy of the files remains outside of the folder that gets updated on Git. 
An additional folder holding datasets is originally found in the shared Dropbox folder `.../Dropbox/Clarkia-LTREB/data and scripts/files from monica/`. I created a folder on my local Dropbox `/Users/Gregor/Dropbox/dataLibrary/` to hold a copy of these files. To copy the files, I run the following in my Terminal:

`cp /Users/Gregor/Dropbox/Clarkia-LTREB/data\ and\ scripts/files\ from\ monica/* /Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/data\ and\ scripts/files\ from\ monica/`

I last ran this on 02/06/20 at 10:50 AM. The files in this directory are now a copy of those in the shared Dropbox folder. For the time being, the copy of the files remains outside of the folder that gets updated on Git. 
Data for seed pots was shared via email on 04/11/2020 as `seed_pot_data_updated.xlsx`.

Additional changes:

- in `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/new fruit & seed files/fruit&seeds_2018.xlsx`, on the first sheet, I deleted cell V141 because it seemed that this was a formula (an average) rather than including raw data.
- in `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/new fruit & seed files/fruit&seeds_2014.xlsx` I changed cell Q254 from `0q` to `0`

Additional changes:

- saved `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/seeds_2011.xls` as `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/seeds_2011.xlsx`
- saved `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/seeds_2012.xls` as `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/seeds_2012.xlsx`
- deleted empty rows at the bottom of the file `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/seeds_2012.xlsx`; these are filled with some formatting and otherwise may give problems with data importing

- saved `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/seeds_2013.xls` as `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/seeds_2013.xlsx`
- deleted empty rows at the bottom of the file `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/seeds_2013.xlsx`; these are filled with some formatting and otherwise may give problems with data importing

- saved `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/seeds_2014.xls` as `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/seeds_2014.xlsx`
- deleted empty rows at the bottom of the file `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/seeds_2014.xlsx`, sheet 2; these are filled with some formatting and otherwise may give problems with data importing

- saved `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/seeds_2015.xls` as `/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/seeds_2015.xlsx`


### Processing data {#processing-data-link}

Load the libraries for data processing (see https://github.com/r-lib/rlang/issues/669 for the overwrite message I am suppressing)

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(knitr)
library(lubridate)
library(tidyxl)
library(readr)
```

Scripts for the original processing I did is located in the following folder. I want to document when I last updated these files before I start updating the workflow. None of the original scripts are going to be changed. 

```{r, message=FALSE, warning=FALSE}
dir = c("/Users/Gregor/Dropbox/projects/clarkiaScripts/code/reshapeDataScripts/")
files = list.files(path=dir)
dt <- data.frame(files)
fileModified <- function(x){file.info(paste0(dir,x))$mtime}

for(i in 1:dim(dt)[1]){
  dt[i,2] <- ymd_hms(fileModified(files[i])) %>%
    data.frame() %>% rename(dateTime='.') 
}

kable(dt)
```

### Seedlings and fruiting plant data

Start with the survival data. This is file `.../reshapeSeeds.R` file in the list above.

```{r, message=FALSE, warning=FALSE}
directory=c("/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/")
# 

# extract data and write out temporary csv
tmp <- readxl::read_excel(paste0(directory,"Survivorship & Fecundity_06-15.xls"), 
                      sheet = 1, na = c("","NA" )) %>%
  janitor::clean_names(case="lower_camel") %>%
  readr::write_csv(paste0("~/Dropbox/dataLibrary/temp/","survivorshipFecundity","-raw.csv"))

names(tmp)

# remove conditional and easting/northing
tmp <- tmp %>%
  dplyr::select(-contains("numberFruitplSdl")) %>%
  dplyr::select(-contains("fruitplNumberSdl")) %>%
  dplyr::select(-c("easting","northing"))

names(tmp)

# remove fruits per plant averages
tmp <- tmp %>%
  dplyr::select(-contains("fruitPl",ignore.case=FALSE)) %>%
  dplyr::select(-contains("undamagedFruitPl")) %>%
  dplyr::select(-contains("damagedFruitPl"))

names(tmp)

# select relevant columns
tmp <- tmp %>%
  dplyr::select(contains(c("site","transect","position","seedlingNumber","fruitplNumber"),ignore.case=FALSE))

names(tmp)
# remove 

censusSeedlingsFruitingPlants <- tmp %>%
  tidyr::pivot_longer(cols=contains(c("seedlingNumber","fruitplNumber")),
               names_to = c("variable","year"),
               names_pattern = "(.*)_(.*)",
               values_to = "count") %>%
  tidyr::separate(variable,into = c("variable", "month"), "(?<=[a-z])(?=[0-9])") %>%
  dplyr::mutate(year = as.numeric(paste0(20,year))) %>%
  dplyr::mutate(position = as.character(position))

censusSeedlingsFruitingPlants <- censusSeedlingsFruitingPlants %>%
  dplyr::select(-month) %>%
  tidyr::pivot_wider(names_from = variable,
                     values_from = count) 

readr::write_rds(censusSeedlingsFruitingPlants,"~/Dropbox/dataLibrary/postProcessingData/censusSeedlingsFruitingPlants.RDS")

```
Here, I summarize the seedling to fruiting plant dataset in two ways. First, I plot the total number of estimates for seedling survival to fruiting per population and year.

```{r, message=FALSE, warning=FALSE}
tmp1 <- censusSeedlingsFruitingPlants %>%
  dplyr::filter(!(seedlingNumber==0&fruitplNumber==0)) %>%
  dplyr::group_by(year,site) %>%
  dplyr::summarise(count = sum(!is.na(seedlingNumber))) %>%
  tidyr::pivot_wider(names_from=year,values_from=count)

kable(tmp1, caption="Summary table of the number of estimates for seedling survival to fruiting")

 #print(xtable::xtable(tmp, type = "latex"), include.rownames=FALSE)
```

Second, I plot the number of estimates for seedling survival to fruiting per population and year that do not have fewer seedlings than fruiting plants

```{r, message=FALSE, warning=FALSE}
tmp <- censusSeedlingsFruitingPlants %>%
  dplyr::filter(!(seedlingNumber==0&fruitplNumber==0)) %>%
  dplyr::group_by(year,site) %>%
  dplyr::filter(fruitplNumber<=seedlingNumber) %>%
  dplyr::summarise(count = sum(!is.na(seedlingNumber))) %>%
  tidyr::pivot_wider(names_from=year,values_from=count)

vals<-signif((1-tmp[,-1]/tmp1[,-1])*100,2)

tmp1 = cbind(tmp[,1],vals)
kable(tmp1, caption="Summary table of the percentage of plots with fruiting plant counts exceeding seedling counts")

# print(xtable::xtable(tmp1, type = "latex"), include.rownames=FALSE)
```

### Fruits per plant data for transects

From 2007-2012, there is data on the number of total fruit equivalents on each plant in each permanent plot. In the Excel files, this means that each year's sheet is a ragged array in which a row corresponds to a permanent plot. Each row has a different number of columns because each plot has a different number of plants, and each plant is recorded in its own cell.

The code below takes these Excel files and converts them into a long data frame with the following columns ...

```{r, message=FALSE, warning=FALSE}
# directory for excel files
directory = "/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/new fruit & seed files/"

# years for datasets
years = 2007:2012

# names of the files
namesFruitPerPlantFiles = list.files(directory)[3:8]

#range of the data in each file
rangeFruitPerPlantColumnNames = c("A1:AL602","A1:BC602",
                                  "A1:AU602","A1:AT602",
                                  "A1:AN602","A1:U602")

# range of the data in each file
rangeFruitPerPlantData = c("A3:AL602","A3:BC602",
                                  "A3:AU602","A3:AT602",
                                  "A3:AN602","A3:U602")

# empty list
listDataFrames20072012 <- list()

# loop to create data frame with each year of data
for(i in 1:length(namesFruitPerPlantFiles)){
  
  # extract names from first row
(cnames <- read_excel(paste0(directory,namesFruitPerPlantFiles[i]), 
                      sheet = 1, range = rangeFruitPerPlantColumnNames[i], 
                      na = "NA", n_max=0) %>% 
    names())

  # extract data and write temporary csv 
tmp <- read_excel(paste0(directory,namesFruitPerPlantFiles[i]), 
                            sheet = 1, range = rangeFruitPerPlantData[i],
              na = "NA",  col_names=cnames) %>%
  janitor::clean_names() %>% readr::write_csv(paste0("~/Dropbox/dataLibrary/temp/fruit&seed_data_",years[i],"-raw.csv"))

# remove columns with summary statistics
tmp <- tmp %>%
  dplyr::select(-contains(c("fruitpl_number","average")) )

# get names of variables
keyVariables = names(tmp)

# get columns with data
tmp2 <- tmp %>%
  dplyr::select(-c(site,transect,position))

# rename columns
plantNumbers <- paste0("plant_",1:dim(tmp2)[2])

# rename data frame
names(tmp) <- c(keyVariables[1:3],plantNumbers)

# use pivot_longer to create long form data
# remove rows with NA (corresponding to no plants)
countFruitsPerPlantFromTransects <- tmp %>%
  tidyr::pivot_longer(cols=contains(c("plant")),
               names_to = "plantNumber",
               values_to = "countFruitsPerPlant") %>%
  dplyr::filter(!is.na(countFruitsPerPlant))
  
# add data frame to list
listDataFrames20072012[[i]] <- countFruitsPerPlantFromTransects

}

# append year to data frames  
for(i in 1:length(years)){
  listDataFrames20072012[[i]] <- listDataFrames20072012[[i]] %>%
      dplyr::mutate(year=years[i])
}

# unlist and bind data frames
countFruitsPerPlantTransects <- listDataFrames20072012 %>%
  purrr::reduce(full_join)

# write data frame to RDS
readr::write_rds(countFruitsPerPlantTransects,"~/Dropbox/dataLibrary/postProcessingData/countFruitsPerPlantTransects.RDS")

```
Summary tables

```{r, message=FALSE, warning=FALSE}
tmp <- countFruitsPerPlantTransects %>%
  dplyr::group_by(year,site) %>%
  dplyr::summarise(count = sum(!is.na(countFruitsPerPlant))) %>%
  tidyr::pivot_wider(names_from=year,values_from=count)

tmp <- arrange(tmp,tmp$site)

kable(tmp, caption="Summary of dataset on total fruit equivalents per plant from transects")

# print(xtable::xtable(tmp, type = "latex"), include.rownames=FALSE)
```

From 2013-2018, there is data on the number of undamaged and damaged fruits on each plant in each permanent plot. In the Excel files, this means that each year's sheet is a ragged array in which a row corresponds to a permanent plot. Each row has a different number of columns because each plot has a different number of plants, and each plant is recorded in its own cell.

The code below takes these Excel files and converts them into a long data frame with the following columns ...

```{r, message=FALSE, warning=FALSE}
# directory for excel files
directory = "/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/new fruit & seed files/"

# years for datasets
years = 2013:2018

# names of the files
namesFruitPerPlantFiles = list.files(directory)[9:14]

#range of the data in each file
rangeUndamagedFruitPerPlantColumnNames = c("A1:M602","A1:M602",
                                  "A1:Q602","A1:M602",
                                  "A1:W602","A1:Y602")

rangeDamagedFruitPerPlantColumnNames = c("N1:T602","N1:T602",
                                         "R1:AB602","N1:T602",
                                         "X1:AN602","AA1:AS602")

# range of the data in each file
rangeUndamagedFruitPerPlantData = c("A3:M602","A3:M602",
                                  "A3:Q602","A3:M602",
                                  "A3:W602","A3:Y602")

rangeDamagedFruitPerPlantData = c("N3:T602","N3:T602",
                                    "R3:AB602","N3:T602",
                                    "X3:AN602","AA3:AS602")

# empty list
listDataFrames20132018Undamaged <- list()
listDataFrames20132018Damaged <- list()

# loop to create data frame with each year of data
for(i in 1:length(namesFruitPerPlantFiles)){
  
  # extract names from first row
(cnames <- read_excel(paste0(directory,namesFruitPerPlantFiles[i]), 
                      sheet = 1, range = rangeUndamagedFruitPerPlantColumnNames[i], 
                      na = "NA", n_max=0) %>% 
    names())

  # extract data and write temporary csv 
tmp <- read_excel(paste0(directory,namesFruitPerPlantFiles[i]), 
                            sheet = 1, range = rangeUndamagedFruitPerPlantData[i],
              na = "NA",  col_names=cnames) %>%
  janitor::clean_names() %>% readr::write_csv(paste0("~/Dropbox/dataLibrary/temp/fruit&seed_data_",years[i],"-undamaged-raw.csv"))

# remove columns with summary statistics
tmp <- tmp %>%
  dplyr::select(-contains(c("fruitpl_number","average","undamaged_fruit_pl","damaged_fruit_pl")) )

# get names of variables
keyVariables = names(tmp)

# get columns with data
tmp2 <- tmp %>%
  dplyr::select(-c(site,transect,position))

# rename columns
plantNumbers <- paste0("plant_",1:dim(tmp2)[2])

# rename data frame
names(tmp) <- c(keyVariables[1:3],plantNumbers)

# use pivot_longer to create long form data
# remove rows with NA (corresponding to no plants)
countUndamagedFruitsPerPlantFromTransects <- tmp %>%
  tidyr::pivot_longer(cols=contains(c("plant")),
               names_to = "plantNumber",
               values_to = "countUndamagedFruitsPerPlant") %>%
  dplyr::filter(!is.na(countUndamagedFruitsPerPlant)) 
  
# add data frame to list
listDataFrames20132018Undamaged[[i]] <- countUndamagedFruitsPerPlantFromTransects

## now do the damaged fruit counts

  # extract names from first row
(cnames <- read_excel(paste0(directory,namesFruitPerPlantFiles[i]), 
                      sheet = 1, range = rangeDamagedFruitPerPlantColumnNames[i], 
                      na = "NA", n_max=0) %>% 
    names())

  # extract data and write temporary csv 
tmpDamaged <- read_excel(paste0(directory,namesFruitPerPlantFiles[i]), 
                            sheet = 1, range = rangeDamagedFruitPerPlantData[i],
              na = "NA",  col_names=cnames) %>%
  janitor::clean_names() %>% readr::write_csv(paste0("~/Dropbox/dataLibrary/temp/fruit&seed_data_",years[i],"-damaged-raw.csv"))

# rename columns
plantNumbersDamaged <- paste0("plant_",1:dim(tmpDamaged)[2])

# rename data frame
names(tmpDamaged) <- plantNumbersDamaged

# add columns with variable names
tmpDamaged <- tmp %>%
  dplyr::select(c(site,transect,position)) %>%
  dplyr::bind_cols(tmpDamaged)

# use pivot_longer to create long form data
# remove rows with NA (corresponding to no plants)
countDamagedFruitsPerPlantFromTransects <- tmpDamaged %>%
  tidyr::pivot_longer(cols=contains(c("plant")),
               names_to = "plantNumber",
               values_to = "countDamagedFruitsPerPlant") %>%
  dplyr::filter(!is.na(countDamagedFruitsPerPlant)) 
  
listDataFrames20132018Damaged[[i]] <- countDamagedFruitsPerPlantFromTransects

}

# append year to data frames  
for(i in 1:length(years)){
  listDataFrames20132018Undamaged[[i]] <- listDataFrames20132018Undamaged[[i]] %>%
      dplyr::mutate(year=years[i],
                    damage=0)
  
    listDataFrames20132018Damaged[[i]] <- listDataFrames20132018Damaged[[i]] %>%
      dplyr::mutate(year=years[i],
                    damage=1)
}

# unlist and bind data frames
countUndamagedFruitsPerPlantTransects <- listDataFrames20132018Undamaged %>%
  purrr::reduce(full_join)

countDamagedFruitsPerPlantTransects <- listDataFrames20132018Damaged %>%
  purrr::reduce(full_join)

countUndamagedDamagedFruitsPerPlantTransects<-left_join(countUndamagedFruitsPerPlantTransects,countDamagedFruitsPerPlantTransects,by=c("site","transect","position","plantNumber","year")) %>%
  dplyr::select(-c("damage.x","damage.y"))

# write data frame to RDS
readr::write_rds(countUndamagedDamagedFruitsPerPlantTransects,"~/Dropbox/dataLibrary/postProcessingData/countUndamagedDamagedFruitsPerPlantTransects.RDS")

```

Summary tables

```{r, message=FALSE, warning=FALSE}
tmp <- countUndamagedDamagedFruitsPerPlantTransects %>%
  dplyr::group_by(year,site) %>%
  dplyr::summarise(count = sum(!is.na(countUndamagedFruitsPerPlant))) %>%
  tidyr::pivot_wider(names_from=year,values_from=count)

tmp<-arrange(tmp,tmp$site)

kable(tmp, caption="Summary of dataset on undamaged and damaged fruits per plant from transects")

# print(xtable::xtable(tmp, type = "latex"), include.rownames=FALSE)
```

### Fruits per plant data extra plots

For 2006, there is no data on fruits per plant associated with permanent plots. From 2007-2012, additional plants outside of permanent plots were sampled to supplement counts of fruits per plant.

```{r, message=FALSE, warning=FALSE}
# directory for excel files
directory = "/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/new fruit & seed files/"

# years for datasets
years = 2006:2012

# names of the files
namesFruitPerPlantFiles = list.files(directory)[2:8]

#range of the data in each file
rangeFruitPerPlantData = c("A1:C4580","A1:C3133",
                           "A1:C3755","A1:C3664",
                           "A1:C5249","A1:C3209",
                           "A1:C1222")

# empty list
listDataFrames20062012 <- list()

# loop to create data frame with each year of data
for(i in 1:length(namesFruitPerPlantFiles)){
  
    # extract data and write temporary csv 
tmp <- readxl::read_excel(paste0(directory,namesFruitPerPlantFiles[i]), 
                            sheet = 2, range = rangeFruitPerPlantData[i],
              na = "NA") %>%
  janitor::clean_names("lower_camel") %>%
  readr::write_csv(paste0("~/Dropbox/dataLibrary/temp/fruit&seed_data_",years[i],"-extra-raw.csv"))

tmp <- tmp %>%
  dplyr::rename(countFruitNumberPerPlant=undamagedFruitNumberPerPlant) %>%
  dplyr::mutate(damage=0)

# add data frame to list
listDataFrames20062012[[i]] <- tmp

}

# append year to data frames  
for(i in 1:length(years)){
  listDataFrames20062012[[i]] <- listDataFrames20062012[[i]] %>%
      dplyr::mutate(year=years[i])
}

# unlist and bind data frames
countFruitsPerPlantAllPlots <- listDataFrames20062012 %>%
  purrr::reduce(full_join)

# write data frame to RDS
readr::write_rds(countFruitsPerPlantAllPlots,"~/Dropbox/dataLibrary/postProcessingData/countFruitsPerPlantAllPlots.RDS")

```

```{r, message=FALSE, warning=FALSE}
tmp <- countFruitsPerPlantAllPlots %>%
  dplyr::filter(permanentPlot==0) %>%
  dplyr::group_by(year,site) %>%
  dplyr::summarise(count = sum(!is.na(countFruitNumberPerPlant))) %>%
  tidyr::pivot_wider(names_from=year,values_from=count)

tmp<-arrange(tmp,tmp$site)

kable(tmp, caption="Summary of dataset on total fruit equivalents per plant from extra plots")

# print(xtable::xtable(tmp, type = "latex"), include.rownames=FALSE)
```


### Seeds per fruit data {#seeds-per-fruit-data-link}

Start with the seed data. This is file `.../reshapeSeeds.R` file in the list above.

The data files for 2006-2010 have two columns. The data files include 1 sheet with the following data: 1 column for the site at which the undamaged fruit was collected, and 1 column for the number of seeds in the undamaged fruit.

```{r, message=FALSE, warning=FALSE}
directory = "/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/new fruit & seed files/"

# all years in the dataset
years <- 2006:2018

# all sheets that contain this data
sheetVector <- c("06 Raw Data","07 Raw Data","08 Raw Data",
                 "09 Raw Data","10 Raw Data","11 Raw Data",
                 "12 Raw Data","13 Raw Data","14 Raw Data",
                 "15 Raw Data","16 Raw Data","17 Raw Data",
                 "18 Raw Data")

# range of data in each sheet
rangeVector <- c("Q5:V828","Q6:V1277","Q4:V1188",
                 "P4:U1214","Q5:V1248","Q8:V1051",
                 "Q7:V654","R12:W988","R10:W1110",
                 "R10:W1077","S10:X1532","R9:W1667",
                 "R12:W1311")

# empty list
dataList <- list()

# loop for getting each data range into a list of data frames
for(i in 1:length(sheetVector)){
  dataList[[i]] <- read_excel(
    paste0(directory,"Above_ground survival & fecundity 06-19.xlsx"),
    sheet = sheetVector[i], 
    range = rangeVector[i],
    na = "NA")
}

# rename variable to be consistent with other datasets
names(dataList[[3]])[6] <- "demography?"

# add year to the data frames
for(i in 1:length(years)){
  dataList[[i]] <- dataList[[i]] %>%
      dplyr::mutate(year=years[i])
}

# join all data frames
countSeedPerFruit <- dataList %>%
  purrr::reduce(full_join)

# clean up variable names and use lower camelcase for variable names
countSeedPerFruit <- countSeedPerFruit %>% 
  janitor::clean_names(case="lower_camel")

# save the dataframe as an RDS file
readr::write_rds(countSeedPerFruit,"~/Dropbox/dataLibrary/postProcessingData/countSeedPerFruit.RDS")

```

Summary table.

```{r, message=FALSE, warning=FALSE}
tmp <- countSeedPerFruit %>%
  dplyr::filter(damaged==0) %>%
  dplyr::filter(demography==1) %>%
  dplyr::group_by(year,site) %>%
  dplyr::summarise(count = sum(!is.na(sdno))) %>%
  tidyr::pivot_wider(names_from=year,values_from=count)

kable(tmp, caption="Summary of dataset on seeds per undamaged fruit")

# print(xtable::xtable(tmp, type = "latex"), include.rownames=FALSE)
```
Summary table.

```{r, message=FALSE, warning=FALSE}
tmp <- countSeedPerFruit %>%
  dplyr::filter(damaged==1) %>%
  dplyr::filter(demography==1) %>%
  dplyr::group_by(year,site) %>%
  dplyr::summarise(count = sum(!is.na(sdno))) %>%
  tidyr::pivot_wider(names_from=year,values_from=count)

kable(tmp, caption="Summary of dataset on seeds per damaged fruit")

# print(xtable::xtable(tmp, type = "latex"), include.rownames=FALSE)
```

The data files for 2011-2012 have four columns: (1) a column for the site at which the undamaged fruit was collected, (2) a column for the number of seeds in the undamaged fruit, and (3) a column indicating whether the fruit was collected from a permanent plot or from across the site. (4) An additional column has notes: some seed counts were randomly resampled from previous years' counts and some sites did not have any undamaged fruits in the given year. 

```{r, message=FALSE, warning=FALSE}
# yearExtract<-function(x){
#   tmp<-as.numeric(sapply(strsplit(x, c("[seeds_.xls]")), "["))
#   tmp<-tmp[!is.na(tmp)]
#   tmp
# }
# 
# years <- unlist(lapply(fileList[6:7],yearExtract))
# 
# reshapeFun <- function(x){
# x %>%
#   dplyr::rename(site='Site') %>%
#   dplyr::rename(seedCount = `Seed no per fruit`) %>%
#   dplyr::mutate(damageStatusBinary = 0)
# }
# 
# appendFontColor <- function(x){
# 
# # Step 1: import the table taking only cell values and ignoring the formatting
# tidyExcel <- read_excel(x,na="NA") 
# tidyExcel <- reshapeFun(tidyExcel)
# 
# # Step 2: import one column of the table, taking only the formatting and not the
# # cell values
# 
# # `formats` is a pallette of fill colours that can be indexed by the
# # `local_format_id` of a given cell to get the fill colour of that cell
# font_colours <- xlsx_formats(x)$local$font$color$rgb
# 
# # Import all the cells, filter out the header row, filter for the first column,
# # and create a new column `font_colour` of the font colours, by looking up the
# # local_format_id of each cell in the `font_colours` pallette.
# fonts <- xlsx_cells(x, sheet = 1) %>%
#   filter(row >= 2, col == 1) %>% # Omit the header row
#   mutate(font_colour = font_colours[local_format_id]) %>%
#   select(font_colour)
# 
# # Step 3: append the `font` column to the rest of the data
# tmp <- bind_cols(tidyExcel, fonts)
# 
# out <- tmp %>% 
#   dplyr::select(site,seedCount,permanentPlot,damageStatusBinary,font_colour) %>%
#   dplyr::rename(permanentPlotBinary=permanentPlot) %>%
#   dplyr::mutate(fieldData = ifelse(font_colour=="FF000000",1,0)) %>%
#   dplyr::select(-font_colour)
# 
# return(out)
# 
# }
# 
# fileDfs<-lapply(fileList[6:7],appendFontColor)
# 
# for(i in 1:length(years)){
#   fileDfs[[i]] <- fileDfs[[i]] %>%
#       dplyr::mutate(year=years[i])
# }
# 
# seed20112012 <- fileDfs %>% 
#   purrr::reduce(full_join)
# 
# seed20112012 %>% 
#   dplyr::filter(permanentPlotBinary!=fieldData)
```

From 2013-2015, we collected data on the number of seeds per undamaged fruit and the number of seeds per damaged fruit. The data files thus include two sheets, 1 for the number of seeds per undamaged fruit and 1 for the number of seeds per damaged fruit. At each site, undamaged fruits were collected from plants spread across the entire site. The data files thus include 1 column for the site at which the undamaged fruit was collected, and 1 column for the number of seeds in the fruit. 

The sheets for undamaged fruits have the following columns: 1 column for the site at which the undamaged fruit was collected, 1 column for the number of seeds in the undamaged fruit, and 1 column indicating whether the fruit was collected from a permanent plot or from across the site. A fourth column has notes: some seed counts were randomly resampled from previous years' counts and some sites did not have any undamaged fruits in the given year. Some of the data files have empty columns between the data and the notes.

```{r, message=FALSE, warning=FALSE}
# yearExtract<-function(x){
#   tmp<-as.numeric(sapply(strsplit(x, c("[seeds_.xls]")), "["))
#   tmp<-tmp[!is.na(tmp)]
#   tmp
# }
# 
# years <- unlist(lapply(fileList[8:10],yearExtract))
# 
# reshapeFun <- function(x){
# x %>%
#   dplyr::rename(site='Site') %>%
#   dplyr::rename(seedCount = `Seed no per undamaged fruit`) %>%
#   dplyr::mutate(damageStatusBinary = 0)
# }
# 
# appendFontColor <- function(x){
# 
# # Step 1: import the table taking only cell values and ignoring the formatting
# tidyExcel <- read_excel(x,na="NA",sheet=1) 
# tidyExcel <- reshapeFun(tidyExcel)
# 
# # Step 2: import one column of the table, taking only the formatting and not the
# # cell values
# 
# # `formats` is a pallette of fill colours that can be indexed by the
# # `local_format_id` of a given cell to get the fill colour of that cell
# font_colours <- xlsx_formats(x)$local$font$color$rgb
# 
# # Import all the cells, filter out the header row, filter for the first column,
# # and create a new column `font_colour` of the font colours, by looking up the
# # local_format_id of each cell in the `font_colours` pallette.
# fonts <- xlsx_cells(x, sheet = 1) %>%
#   filter(row >= 2, col == 1) %>% # Omit the header row
#   mutate(font_colour = font_colours[local_format_id]) %>%
#   select(font_colour)
# 
# # Step 3: append the `font` column to the rest of the data
# tmp <- bind_cols(tidyExcel, fonts)
# 
# out <- tmp %>% 
#   dplyr::select(site,seedCount,permanentPlot,damageStatusBinary,font_colour) %>%
#   dplyr::rename(permanentPlotBinary=permanentPlot) %>%
#   dplyr::mutate(fieldData = ifelse(font_colour=="FF000000"|font_colour=="FF008000",1,0)) %>%
#   dplyr::select(-font_colour)
# 
# return(out)
# 
# }
# 
# fileDfs<-lapply(fileList[8:10],appendFontColor)
# 
# for(i in 1:length(years)){
#   fileDfs[[i]] <- fileDfs[[i]] %>%
#       dplyr::mutate(year=years[i])
# }
# 
# seedUndamaged20132015 <- fileDfs %>% 
#   purrr::reduce(full_join) 
# 
# seedUndamaged20132015 %>% 
#   dplyr::filter(permanentPlotBinary!=fieldData)
```


The sheets for damaged fruits have the following columns: 1 column for the site at which the damaged fruit was collected, and 1 column for the number of seeds in the damaged fruit. Another column has notes: some sites did not have any damaged fruits in the given year. Some of the data files have empty columns between the data and the notes.

```{r, message=FALSE, warning=FALSE}
# yearExtract<-function(x){
#   tmp<-as.numeric(sapply(strsplit(x, c("[seeds_.xls]")), "["))
#   tmp<-tmp[!is.na(tmp)]
#   tmp
# }
# 
# years <- unlist(lapply(fileList[8:10],yearExtract))
# 
# reshapeFun <- function(x){
# x %>%
#   dplyr::rename(site='Site') %>%
#   dplyr::rename(seedCount = `Seed no per damaged fruit`) %>%
#   dplyr::mutate(damageStatusBinary = 1, permanentPlot = NA)
# }
# 
# appendFontColor <- function(x){
# 
# # Step 1: import the table taking only cell values and ignoring the formatting
# tidyExcel <- read_excel(x,na=c("NA","MA"),sheet=2) 
# tidyExcel <- reshapeFun(tidyExcel)
# 
# # Step 2: import one column of the table, taking only the formatting and not the
# # cell values
# 
# # `formats` is a pallette of fill colours that can be indexed by the
# # `local_format_id` of a given cell to get the fill colour of that cell
# font_colours <- xlsx_formats(x)$local$font$color$rgb
# 
# # Import all the cells, filter out the header row, filter for the first column,
# # and create a new column `font_colour` of the font colours, by looking up the
# # local_format_id of each cell in the `font_colours` pallette.
# fonts <- xlsx_cells(x, sheet = 2) %>%
#   filter(row >= 2, col == 1) %>% # Omit the header row
#   mutate(font_colour = font_colours[local_format_id]) %>%
#   select(font_colour)
# 
# # Step 3: append the `font` column to the rest of the data
# tmp <- bind_cols(tidyExcel, fonts)
# 
# out <- tmp %>% 
#   dplyr::select(site,seedCount,permanentPlot,damageStatusBinary,font_colour) %>%
#   dplyr::rename(permanentPlotBinary = permanentPlot) %>%
#   dplyr::mutate(fieldData = ifelse(font_colour=="FF000000"|font_colour=="FF008000",1,0)) %>%
#   dplyr::select(-font_colour)
# 
# return(out)
# 
# }
# 
# fileDfs<-lapply(fileList[8:10],appendFontColor)
# 
# for(i in 1:length(years)){
#   fileDfs[[i]] <- fileDfs[[i]] %>%
#       dplyr::mutate(year=years[i])
# }
# 
# seedDamaged20132015 <- fileDfs %>% 
#   purrr::reduce(full_join) 
# 
# seedDamaged20132015 %>% 
#   dplyr::filter(permanentPlotBinary!=fieldData)
```

It might be useful to summarize the datasets:

```{r, message=FALSE, warning=FALSE}
# need to rewrite this code block

# summary20062010<-seed20062010 %>%
#   dplyr::group_by(year,site) %>%
#   dplyr::summarise(count = sum(!is.na(seedCount))) %>%
#   tidyr::spread(key="year",value="count")
# 
# summary20112012<-seed20112012 %>%
#   dplyr::group_by(year,site) %>%
#   dplyr::summarise(count = sum(!is.na(seedCount))) %>%
#   tidyr::spread(key="year",value="count")
# 
# summaryUndamaged20132015<-seedUndamaged20132015  %>%
#   dplyr::group_by(year,site) %>%
#   dplyr::summarise(count = sum(!is.na(seedCount))) %>%
#   tidyr::spread(key="year",value="count")
# 
# summaryDamaged20132015<-seedDamaged20132015  %>%
#   dplyr::group_by(year,site) %>%
#   dplyr::summarise(count = sum(!is.na(seedCount))) %>%
#   tidyr::spread(key="year",value="count")
# 
# summaryTable <- summary20062010 %>%
#   dplyr::full_join(summary20112012) %>%
#   dplyr::full_join(summaryUndamaged20132015)
# 
# kable(summaryTable, caption="Summary table of the number of counts of seeds from undamaged fruits")
# 
# kable(summaryDamaged20132015, caption="Summary table of the number of counts of seeds from damaged fruits")
```

### Seed bag data

Data for seed bags from the field.

```{r, message=FALSE, warning=FALSE}
dir=c("/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/20_demography_sites/")

tmp <- readxl::read_excel(paste0(dir,"seed bag data.xls"),sheet = 1,na=c("NA","?"))

seedBags <- tmp %>%
  dplyr::rename(bagNo = 'bag #') %>%
  dplyr::rename(round = 'Round') %>%
  dplyr::rename(yearStart = 'Start_yr(Oct)') %>%
  dplyr::rename(age = 'Age') %>%
  dplyr::rename(yearData = 'Yr_germ(Jan)/viability(Oct)') %>%
  dplyr::rename(seedlingJan = 'Jan_germ') %>%
  dplyr::rename(intactJan = 'Jan_intact') %>%
  dplyr::rename(totalJan = 'Jan_total') %>%
  dplyr::rename(intactOct = 'Oct_intact') %>%
  dplyr::select(-c("viability%"))

seedBagsData <- seedBags

readr::write_rds(seedBagsData,"~/Dropbox/dataLibrary/postProcessingData/seedBagsData.rds")

# january germinants
summarySeedBags<-seedBags  %>%
  tidyr::unite("yearAge", yearData,age,sep="-age") %>%
  dplyr::group_by(yearAge,site) %>%
  dplyr::summarise(count = sum(!is.na(seedlingJan))) %>%
  tidyr::spread(key=c("yearAge"),value="count")

kable(summarySeedBags, caption="Summary table of the number of seed bags counted in January for intact and germinated seeds")

# october intacts
summarySeedBags<-seedBags  %>%
  tidyr::unite("yearAge", yearData,age,sep="-age") %>%
  dplyr::group_by(yearAge,site) %>%
  dplyr::summarise(count = sum(!is.na(intactOct))) %>%
  tidyr::spread(key=c("yearAge"),value="count")

kable(summarySeedBags, caption="Summary table of the number of seed bags counted in October for intact seeds")

```

### Viability trial data

```{r, message=FALSE, warning=FALSE}
dir=c("/Users/Gregor/Dropbox/dataLibrary/Clarkia-LTREB/data and scripts/files from monica/")

tmp <- read_excel(paste0(dir,"germ_viab_data_final_updated.xlsx"),sheet = 1,na=c("NA","?","."))

viabilityRawData <- tmp %>%
  dplyr::rename(site = 'population') %>%
  dplyr::rename(bagNo = 'bag#') %>%
  dplyr::rename(round = 'Round') %>%
  dplyr::rename(age = 'Age') %>%
  dplyr::rename(block = 'Block') %>%
  dplyr::rename(germStart = '#tested_for_germination') %>%
  dplyr::rename(germCount = '#germinating') %>%
  dplyr::rename(germPerc = '%Germ') %>%
  dplyr::rename(germNot = '#not germinated') %>%
  dplyr::rename(viabStart = '#tested_for_viability') %>%
  dplyr::rename(viabStain = '#viable(stained_red)') %>%
  dplyr::rename(viabPerc = '%viability') %>%
  dplyr::rename(condTest = '#not_germinated=#tested_for_viability?') %>%
  dplyr::rename(viabPerc2 = '%Viabil') %>%
  dplyr::rename(viabTotal = 'Overall_viability')

readr::write_rds(viabilityRawData,"~/Dropbox/dataLibrary/postProcessingData/viabilityRawData.rds")

# germination trial 
summaryGerminationTrials<-viabilityRawData  %>%
  tidyr::unite("roundAge", round,age,sep="-age") %>%
  dplyr::group_by(roundAge,site) %>%
  dplyr::summarise(count = sum(germStart,na.rm=TRUE)) %>%
  # use following line to count number of trials
  # dplyr::summarise(count = sum(!is.na(germStart))) %>%
  tidyr::spread(key=c("roundAge"),value="count")

kable(summaryGerminationTrials, caption="Summary table of the number of germination trials")

# viability trial 
summaryViabilityTrials<-viabilityRawData  %>%
  tidyr::unite("roundAge", round,age,sep="-age") %>%
  dplyr::group_by(roundAge,site) %>%
  dplyr::summarise(count = sum(viabStart,na.rm=TRUE)) %>%
  # use following line to count number of trials
  # dplyr::summarise(count = sum(!is.na(viabStart))) %>%  
  tidyr::spread(key=c("roundAge"),value="count")

kable(summaryViabilityTrials, caption="Summary table of the number of germination trials")

```
Rows of concern in the viability trials. 

```{r, message=FALSE, warning=FALSE}

# rows where the number of seeds that didn't germinate
# is less than the number of seeds that started the viability trials
error<-viabilityRawData %>% dplyr::filter(germNot<viabStart) 

nNotGermLessThanViab = nrow(error)

# add rows where bag lacks number
error <- viabilityRawData %>% dplyr::filter(bagNo=="?") %>% 
  bind_rows(error)

nNotBagNumber = nrow(error) - nNotGermLessThanViab

# add one row where viabStart is NA
# how is this different from no seeds starting
error <- viabilityRawData %>% dplyr::filter(is.na(viabStart)) %>%
  bind_rows(error)

nNotRecordedViabStart = nrow(error) - nNotBagNumber

# conditional test is coded incorrectly
error<-viabilityRawData %>% dplyr::filter(germNot==viabStart&condTest=="N") %>%
  bind_rows(error)

nConditionalTestCoding = nrow(error) - nNotRecordedViabStart


kable(error, caption="Table of rows in the viability trials data set that need to be checked for problems")

```

Possible concerns include:

- the number of seeds that didn't germinate is less than the number of seeds that started the viability trials (`r nNotGermLessThanViab` rows in dataset)
- bags are missing a bag number (`r nNotBagNumber` rows in dataset)
- data missing a number for number of seeds starting viability trials (`r nNotRecordedViabStart` rows in dataset)
- conditional test is coded incorrectly in original data file (`r nConditionalTestCoding` rows in dataset)

The majority of bags had 1-2 tests (equally split). A small number had 3 test and an even smaller had 4. But clearly need to consider that bags were tested twice. 

Data for viability trials in the lab. 

### Seed pot data

```{r, message=FALSE, warning=FALSE}
dir=c("/Users/Gregor/Dropbox/dataLibrary/seedPots/")

tmp <- readxl::read_excel(paste0(dir,"seed_pot_data_updated.xlsx"),sheet = 1,na=c("NA","?","."))

tmp <- tmp %>%
  dplyr::rename(site = 'Site') %>%
  dplyr::select(-Pop) %>%
  dplyr::rename(block = 'Block') %>%
  dplyr::rename(row = 'Row') %>%
  dplyr::rename(column = 'Col') %>%
  dplyr::rename(seedNumber = 'sd#') %>% 
  dplyr::mutate(seedNumber = as.numeric(ifelse(seedNumber=="C",0,seedNumber))) %>%
  dplyr::rename(yearPlaced = 'yr_pl') %>%
  dplyr::rename(yearCollected = 'yr_coll') %>%
  dplyr::rename(yearCensusOne = 'census1') %>%
  dplyr::rename(seedlingNumberCensusOne = 'sdl#_yr1') %>%
  dplyr::rename(yearCensusTwo = 'census2') %>%
  dplyr::rename(seedlingNumberCensusTwo = 'sdl#_yr2') %>%
  dplyr::rename(yearCensusThree = 'census3') %>%
  dplyr::rename(seedlingNumberCensusThree = 'sdl#_yr3')

seedPotData <- tmp

readr::write_rds(seedPotData,"~/Dropbox/dataLibrary/postProcessingData/seedPotData.rds")
```