---
output: github_document
---

# Model development

Here, I document progress developing models for the Clarkia demography project. 

- [Table of datasets](#table-of-datasets)
- [Belowground vital rates](#belowground-vital-rates)
- [Seedling survival to fruiting](#seedling-survival-to-fruiting)
- [Fruits per plant](#fruits-per-plant)
- [Seeds per fruit](#seeds-per-fruit)
- [Transition to seed bank](#transition-to-seed-bank)

I used pandoc to convert the latex file to a md file:

`pandoc -s parameter-table.tex -o parameter-table.md`

# Packages

I am using the packages `tidyverse` and `readxl` (documentation: https://readxl.tidyverse.org/). 

Load the libraries for data processing (see https://github.com/r-lib/rlang/issues/669 for the overwrite message I am suppressing).

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(ggplot2)
```

For model building I am using:

```{r, message=FALSE, warning=FALSE}
library(rjags)
library(tidybayes)
library(MCMCvis)
```

# Table of datasets

The table below summarizes the data generated in this study (also available [as a pdf at this link](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/products/tables/data-summary.pdf)).

  ---------------------------------------- ------------------ ------------------------ -----------
  **Seed vital rates**                      ---                          ---                ---
  Seed survival and germination            Seed bag burial      Y<sub>1</sub>            2006-2009
  Seed viability                           Viability trials    Y<sub>2</sub>             2006-2009
  Seed survival and germination            Seed pots            Y<sub>3</sub>            2013-2019
  **Seedling survival**                   ---                          ---                ---
  Seedling survival to fruiting            Field surveys        Y<sub>4</sub>            2006-2019
  **Fruits per plant**                       ---                          ---                ---
  Total fruit equivalents per plant        Field surveys       Y<sub>5</sub>             2006-2012
  Undamaged and damaged fruits per plant   Field surveys        Y<sub>6</sub>            2013-2019
  Total fruit equivalents per plant        Extra plots          Y<sub>7</sub>            2006-2012
  Undamaged and damaged fruits per plant   Extra plots          Y<sub>8</sub>            2013-2019
  **Seeds per fruit**                       ---                          ---                ---
  Seeds per undamaged fruit                Lab counts           Y<sub>9</sub>            2006-2019
  Seeds per damaged fruit                  Lab counts          Y<sub>0</sub>             2013-2019
  ---------------------------------------- ------------------ ------------------------ -----------

Links to  all the datasets follow:

* [Seed bag burial](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/library/dataProcessingWorkflow.md#seed-bag-data)
* [Viability trials](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/library/dataProcessingWorkflow.md#viability-trial-data)
* [Seed pots](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/library/dataProcessingWorkflow.md#seed-pot-data) - needs to be updated
* [Field surveys for seedling survival to fruiting](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/library/dataProcessingWorkflow.md#seedlings-and-fruiting-plant-data)
* [Field surveys for total fruit equivalents per plant, and undamaged/damaged fruits per plant](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/library/dataProcessingWorkflow.md#fruits-per-plant-data-extra-plots)
* [Lab counts for seeds per fruit from undamaged and damaged fruits](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/library/dataProcessingWorkflow.md#seeds-per-fruit-data)

# Analysis workflow

* Load data
* Rename variables for consistency
* Filter dataset (e.g. in seedling survival model) until I develop measurement model
* Create data object (list) for analysis in JAGS using for e.g. tidybayes
** this has its advantages but need to make sure that variable designations are consistent 
* Initialize chains 
** currently this is done by hand but I think I need to improve on it
** Rather than writing list(list(...),list(...),list(...)) each with a set of parameters, I want to write functions to generate the appropriate initial conditions and then name the inits after they are in lists using a function (or something similar)
* Set conditions for JAGS
* Set parameters to sample
* Initialize chains
* Update chains
* Sample chains
* Save samples, original data, and data object for JAGS

Even just writing out these steps I think there is more organization to be done. For example, I think it's possible to separate the data construction steps from the model fits. So, for example, I could use this file to process the data for analysis in JAGS using a consistent set of principles. I would then save the objects as data lists to a folder, which would then be loaded in the scripts with the model. 

It would also be great to write R files that are read into here rather than copying over the JAGS code, which is liable to change. 

# Belowground vital rates: age 1

## Data organization

### Seed bag data
Begin by importing the seed bag data (Y<sub>1</sub>).

```{r}
seedBagsData <- readRDS("~/Dropbox/dataLibrary/postProcessingData/seedBagsData.rds")
```

The next step is to clean and organize the datasets. Begin by working with the seed bag data.  

* I exclude rows with missing data in January. Bags that were not recovered in January but were recovered in October; we have no way of getting an estimate of how many seeds were present halfway through the trial.
* I exclude the row for ``MC III 5 13 1` which has 91 intact seeds and 17 seedlings (this should not be possible if 100 seeds started the trials).
* I exclude rows with missing data in October.
* I exclude rows where the number of intact seeds in January is less than the number of intact seeds in October.

The filtering steps in the following code chunk remove 64 rows from the dataset. 

```{r}
datasetSize=nrow(seedBagsData)

seedBagsData<-subset(seedBagsData,!is.na(seedBagsData$totalJan))
datasetSize <- cbind(datasetSize,nrow(seedBagsData))

seedBagsData <- seedBagsData %>%
  dplyr::filter(totalJan<=100)
datasetSize <- cbind(datasetSize,nrow(seedBagsData))

seedBagsData<-subset(seedBagsData,!is.na(seedBagsData$intactOct))
datasetSize <- cbind(datasetSize,nrow(seedBagsData))

seedBagsData<-subset(seedBagsData,!(seedBagsData$intactJan<seedBagsData$intactOct))
datasetSize <- cbind(datasetSize,nrow(seedBagsData))
```

I then create a variable for the number of seeds that start the trial.

```{r}
# Create a variable for the number of seeds at the start of the trial
seedBagsData$seedStart<-as.double(100)
```

### Viability trial data

Import the viability trial data (Y<sub>2</sub>)

```{r}
viabilityRawData <- readRDS("~/Dropbox/dataLibrary/postProcessingData/viabilityRawData.rds")
```

Remove summary variables from the dataset.

```{r}
viabilityRawData <- viabilityRawData %>% 
  dplyr::select(-c(germPerc,germNot,viabPerc,viabPerc2,condTest))
```

Create the variable `bag` ()
```{r}
viabilityRawData$bag <-as.integer(as.numeric(viabilityRawData$bagNo))
```
One of the rows in the viability dataset is coded differently: the variable `viabStart` is `NA` instead of `0` as in the rest of the dataset. Correct this here by recoding that value.

```{r}
viabilityRawData[is.na(viabilityRawData$viabStart),]$viabStart = 0
```

I also remove data with any NAs in the viability stain column.

Check the data. Are there any rows where there are more germinants and seeds that start the viability trials, than seeds that start the germination trials? The answer is yes; these rows need to be checked in the data binders. 

```{r}
viabilityRawData %>% dplyr::filter(germStart - germCount - viabStart<0) 

```
I exclude the rows with this issue from the analysis.

```{r}
viabilityRawData<-viabilityRawData %>% 
  dplyr::filter(germStart - germCount - viabStart >= 0)
```

## Indexing for JAGS

First, I filter the dataset to include only 1-year old seeds

```{r}
filterData<-function(x, ageNumber) {
  x %>%
    dplyr::filter(age==ageNumber)
}

seedBagsData1<-filterData(seedBagsData,ageNumber=1)
viabilityRawData1<-filterData(viabilityRawData,ageNumber=1)

seedBagsData2<-filterData(seedBagsData,ageNumber=2)
viabilityRawData2<-filterData(viabilityRawData,ageNumber=2)
```

I assign variables that combine the site, bag number, experimental round, and age. This creates a unique ID for each bag, for each dataset. That unique identifier links the bags across the seed bag and viability trial dataset.

```{r}
seedBagsData1<-seedBagsData1 %>%
  tidyr::unite(col='id', c(site,bagNo,round,age), sep="-", remove=FALSE) %>%
  tidyr::unite(col='siteBag', c(site,bagNo), sep="-", remove=FALSE) %>%
  dplyr::mutate(siteBag = as.factor(siteBag)) 

viabilityRawData1<-viabilityRawData1 %>%
  tidyr::unite(col='id', c(site,bagNo,round,age), sep="-", remove=FALSE) %>%
  tidyr::unite(col='siteBag', c(site,bagNo), sep="-", remove=FALSE) %>%
  dplyr::mutate(siteBag = as.factor(siteBag)) 
```

I create a reference table that joins the seed bag and viability trial data via a unique identifier.

```{r}
referenceTable<-data.frame(id=union(seedBagsData1$id, viabilityRawData1$id)) %>%
  dplyr::mutate(idNo = 1:length(id)) 
```

I then append the identifiers to each dataset.

```{r}
seedBagsData1<-seedBagsData1 %>%
  dplyr::left_join(referenceTable,by="id")

viabilityRawData1<-viabilityRawData1 %>%
  dplyr::left_join(referenceTable,by="id")
```

## Data for JAGS

Variable names need to be consistent in both datasets. I make the following changes:

* Set the `year` variable (`yearStart` in seed bags, `round` in viability trials) to be a factor
* Select variables used in the model fitting
* Rename variables to uniquely refer to each dataset (e.g. `site` renamed as `siteBags` in seed bag dataset)

I also summarize the viability trial dataset. Specifically, I sum each count (e.g. `germStart`) across all replicates from a bag.

```{r}

seedBagsData1 = seedBagsData1 %>%
  dplyr::mutate(year = as.factor(yearStart),
                age = as.factor(age)) %>%
  dplyr::select(site,year,age,totalJan,seedStart,seedlingJan,intactOct) %>%
  dplyr::rename(siteBags = site,
                yearBags = year,
                ageBags = age)

viabilityRawData1 = viabilityRawData1 %>%
  dplyr::mutate(year = as.factor(round),
                age = as.factor(age)) %>%
  dplyr::select(site, year, age, germStart, germCount, viabStart, viabStain, idNo) %>%
  dplyr::rename(siteViab = site,
                yearViab = year,
                ageViab = age,
                bag = idNo) %>%
 # dplyr::mutate(bag = as.factor(bag)) %>%
  dplyr::mutate(germStart = ifelse(is.na(germCount), NA, germStart),
                viabStart = ifelse(is.na(viabStain), NA, viabStart)) %>%
  dplyr::group_by(siteViab,yearViab, ageViab, bag) %>%
  # sum observations in each bag; this is ignoring some variation
  dplyr::summarise(germStart = sum(germStart),
                   germCount = sum(germCount),
                   viabStart = sum(viabStart),
                   viabStain = sum(viabStain))

```

Repeat for age 2 data:

```{r}
seedBagsData2<-seedBagsData2 %>%
  tidyr::unite(col='id', c(site,bagNo,round,age), sep="-", remove=FALSE) %>%
  tidyr::unite(col='siteBag', c(site,bagNo), sep="-", remove=FALSE) %>%
  dplyr::mutate(siteBag = as.factor(siteBag)) 

viabilityRawData2<-viabilityRawData2 %>%
  tidyr::unite(col='id', c(site,bagNo,round,age), sep="-", remove=FALSE) %>%
  tidyr::unite(col='siteBag', c(site,bagNo), sep="-", remove=FALSE) %>%
  dplyr::mutate(siteBag = as.factor(siteBag)) 

referenceTable<-data.frame(id=union(seedBagsData2$id, viabilityRawData2$id)) %>%
  dplyr::mutate(idNo = 1:length(id)) 

seedBagsData2<-seedBagsData2 %>%
  dplyr::left_join(referenceTable,by="id")

viabilityRawData2<-viabilityRawData2 %>%
  dplyr::left_join(referenceTable,by="id")

seedBagsData2 = seedBagsData2 %>%
  dplyr::mutate(year = as.factor(yearStart),
                age = as.factor(age)) %>%
  dplyr::select(site,year,age,totalJan,seedStart,seedlingJan,intactOct) %>%
  dplyr::rename(siteBags = site,
                yearBags = year,
                ageBags = age)

viabilityRawData2 = viabilityRawData2 %>%
  dplyr::mutate(year = as.factor(round),
                age = as.factor(age)) %>%
  dplyr::select(site, year, age, germStart, germCount, viabStart, viabStain, idNo) %>%
  dplyr::rename(siteViab = site,
                yearViab = year,
                ageViab = age,
                bag = idNo) %>%
  #dplyr::mutate(bag = as.factor(bag)) %>%
  dplyr::group_by(siteViab,yearViab, ageViab, bag) %>%
    dplyr::mutate(germStart = ifelse(is.na(germCount), NA, germStart),
                viabStart = ifelse(is.na(viabStain), NA, viabStart)) %>%
  # sum observations in each bag; this is ignoring some variation
  dplyr::summarise(germStart = sum(germStart),
                   germCount = sum(germCount),
                   viabStart = sum(viabStart),
                   viabStain = sum(viabStain))

names(seedBagsData2)=paste(names(seedBagsData2),"2",sep="")
names(viabilityRawData2)=paste(names(viabilityRawData2),"2",sep="")

```
## Pass to JAGS

I use `tiybayes::compose_data` to create a data list that I will pass to JAGS. I reset the number of samples `n` to match the dimension of the seed bag dataset.

Try to only use complete rows.

```{r}
viabilityRawData1<-viabilityRawData1[complete.cases(viabilityRawData1),] %>% dplyr::mutate(bag = as.factor(bag))
viabilityRawData2<-viabilityRawData2[complete.cases(viabilityRawData2),] %>% dplyr::mutate(bag2 = as.factor(bag2))

data <- tidybayes::compose_data(seedBagsData1,viabilityRawData1,
                                seedBagsData2,viabilityRawData2)

data$n = dim(seedBagsData1)[1]
data$n2 = dim(seedBagsData2)[1]

```

## Save for JAGS

```{r}
fileDirectory<- c("/Users/Gregor/Dropbox/dataLibrary/workflow/data/")

saveRDS(data,file=paste0(fileDirectory,"belowgroundDataAgeOneTwo.rds"))

# fileDirectory<- c("/Users/Gregor/Dropbox/dataLibrary/workflow/tidyData/")
# 
# saveRDS(seedBagsData,file=paste0(fileDirectory,"seedBagExperiment.rds"))
# saveRDS(viabilityRawData,file=paste0(fileDirectory,"viabilityExperiment.rds"))
```

Use Y<sub>1</sub> and Y<sub>2</sub> to estimate belowground vital rates. Refer to the [appendix on using conditional probability](https://github.com/gregor-fausto/clarkiaSeedBanks/blob/master/products/manuscript/appendix-x-conditional-probability.pdf).

These data are processed and analyzed in `clarkiaSeedBanks/modelBuild/seedBurial/fullModelScripts.R`.

The JAGS script for this data is in `clarkiaSeedBanks/modelBuild/jagsScriptsSeedBags/hierarchicalLogitCentered.R`.

Basic process. Use the data from seed bag burial experiment, germination and viability trials to calculate the parameter theta. Each experiment is a trial with a binomial likelihood, priors and hyperpriors that are centered and logit transformed. I then take these probabilities use them to derive the logit-transformed probabilities, which I then transform into the probabilities of viability, survival, and germination.

```
model { 

### Hyperpriors for germination and viability trials
for(k in 1:n_siteViab){

  # germination
  mu0_g[k] ~  dnorm(0, 0.001)
  sigma0_g[k] ~ dunif(0,1.5)
  tau0_g[k] <- 1/(sigma0_g[k]*sigma0_g[k])

  # viability
  mu0_v[k] ~  dnorm(0, 0.001)
  sigma0_v[k] ~ dunif(0,1.5)
  tau0_v[k] <- 1/(sigma0_v[k]*sigma0_v[k])
    
    ## Priors for germination and viability trials
    for(i in 1:n_yearViab){

    # germination
    mu_g[k,i] ~ dnorm(mu0_g[k], tau0_g[k])
    sigma_g[k,i] ~ dunif(0,1.5)
    tau_g[k,i] <- 1/(sigma_g[k,i]*sigma_g[k,i])

    # viability
    mu_v[k,i] ~ dnorm(mu0_v[k], tau0_v[k])
    sigma_v[k,i] ~ dunif(0,1.5)
    tau_v[k,i] <- 1/(sigma_v[k,i]*sigma_v[k,i])
    }
  }

### Hyperpriors for seed bag burial
for(k in 1:n_siteBags){

  # theta 1
  mu0_1[k] ~  dnorm(0, 0.001)
  sigma0_1[k] ~ dunif(0,1.5)
  tau0_1[k] <- 1/(sigma0_1[k]*sigma0_1[k])

  # theta 2
  mu0_2[k] ~  dnorm(0, 0.001)
  sigma0_2[k] ~ dunif(0,1.5)
  tau0_2[k] <- 1/(sigma0_2[k]*sigma0_2[k])

  # theta 3
  mu0_3[k] ~  dnorm(0, 0.001)
  sigma0_3[k] ~ dunif(0,1.5)
  tau0_3[k] <- 1/(sigma0_3[k]*sigma0_3[k])

    for(i in 1:n_yearBags){

    # theta 1
    mu_1[k,i] ~ dnorm(mu0_1[k], tau0_1[k])
    sigma_1[k,i] ~ dunif(0,1.5)
    tau_1[k,i] <- 1/(sigma_1[k,i]*sigma_1[k,i])

    # theta 2
    mu_2[k,i] ~ dnorm(mu0_2[k], tau0_2[k])
    sigma_2[k,i] ~ dunif(0,1.5)
    tau_2[k,i] <- 1/(sigma_2[k,i]*sigma_2[k,i])

    # theta 3
    mu_3[k,i] ~ dnorm(mu0_3[k], tau0_3[k])
    sigma_3[k,i] ~ dunif(0,1.5)
    tau_3[k,i] <- 1/(sigma_3[k,i]*sigma_3[k,i])
    }
  }

### Likelihood

## viability trials
for(i in 1:n_bag){

  # alpha
  alpha_g[i] ~ dnorm(mu_g[siteViab[i],yearViab[i]],tau_g[siteViab[i],yearViab[i]])
  alpha_v[i] ~ dnorm(mu_v[siteViab[i],yearViab[i]],tau_v[siteViab[i],yearViab[i]])

  # logit 
  logit(theta_g[i]) <- alpha_g[i]
  logit(theta_v[i]) <- alpha_v[i]

  germCount[i] ~ dbinom( theta_g[i] , germStart[i] )
  viabStain[i] ~ dbinom( theta_v[i] , viabStart[i] )

}

## seed burial experiment
for(i in 1:n){

  # alpha
  alpha_1[i] ~ dnorm(mu_1[siteBags[i],yearBags[i]],tau_1[siteBags[i],yearBags[i]])
  alpha_2[i] ~ dnorm(mu_2[siteBags[i],yearBags[i]],tau_2[siteBags[i],yearBags[i]])
  alpha_3[i] ~ dnorm(mu_3[siteBags[i],yearBags[i]],tau_3[siteBags[i],yearBags[i]])

  # logit 
  logit(theta_1[i]) <- alpha_1[i]
  logit(theta_2[i]) <- alpha_2[i]
  logit(theta_3[i]) <- alpha_3[i]

  # use this as the number of trials for theta_3
  intactJan[i] = totalJan[i]-seedlingJan[i]

  totalJan[i] ~ dbinom(theta_1[i], seedStart[i])
  seedlingJan[i] ~ dbinom(theta_2[i], totalJan[i])
  intactOct[i] ~ dbinom(theta_3[i], intactJan[i])

}

### Derived quantities
for(i in 1:n_siteBags){

  p.i_g[i] ~ dnorm(mu0_g[i],tau0_g[i])
  logit(p_g[i]) <- p.i_g[i]

  p.i_v[i] ~ dnorm(mu0_v[i],tau0_v[i])
  logit(p_v[i]) <- p.i_v[i]

  # Probability of viability
  nu_1[i] = p_g[i] + p_v[i]*(1-p_g[i])

  p.i_1[i] ~ dnorm(mu0_1[i],tau0_1[i])
  logit(p_1[i]) <- p.i_1[i]

  p.i_2[i] ~ dnorm(mu0_2[i],tau0_2[i])
  logit(p_2[i]) <- p.i_2[i]

  p.i_3[i] ~ dnorm(mu0_3[i],tau0_3[i])
  logit(p_3[i]) <- p.i_3[i]

  # Probability of winter survival
  s1[i] = p_1[i]*(p_2[i] + (1-p_2[i])*(nu_1[i])^(1/3))

  # Probability of germination
  g1[i] = p_2[i]/(1-(1-(nu_1[i]^(1/3)))*(1-p_2[i]))

  # Probability of spring and summer survival
  s2[i] = p_3[i]*(nu_1[i]^(2/3))
}

}
```

### Seedling survival to fruiting

I use Y<sub>4</sub> to estimate the probability of seedling survival to fruiting. 

These data are processed and analyzed in `clarkiaSeedBanks/modelBuild/seedlingSurvival/fullModelScripts.R`.

The JAGS script for this data is in `clarkiaSeedBanks/modelBuild/jagsScriptsSeedlingSurvival/hierarchicalLogitCentered.R`.

Explore results using different ways of dealing with undercounting.

This is how I calculated the probability of survival to fruiting in fall 2019. I think this data is amenable to the same approach as the one above. Rather than  using hyperpriors at the site level, I will only get per-year estimates for all datasets. 

From `seedBurial/fullModelScripts`:
```
model { 
# hyperpriors
for(i in 1:nsites){
alphaS[i] ~ dnorm(0, .001)
}

for(i in 1:nyears){
betaS[i] ~ dnorm(0, .001)
}

for(j in 1:nsites){
for(k in 1:nyears){
gammaS[j,k] ~ dnorm(0, 0.001)
}
}


# likelihoods
for (i in 1:N){ 
phi[i] = ilogit(alphaS[siteSigma[i]] + betaS[yearSigma[i]] + gammaS[siteSigma[i],yearSigma[i]])
y[i] ~ dbinom(phi[i], n[i])

# simulated data for posterior predictive checks
y.sim[i] ~ dbinom(phi[i], n[i]) 
} 

}
```

### Fruits per plant 

Use Y<sub>5</sub> and Y<sub>7</sub> to estimate the number of fruits per plant.

These data are processed and analyzed in `clarkiaSeedBanks/modelBuild/fecundity/fitness_model.R`.

The JAGS script for this data is in `clarkiaSeedBanks/modelBuild/jagsScriptsFecundity/fecJags.R`.

Explore how to combine data from permanent plots vs. all plots at the site. Perhaps for the fruits per plant estimate use data from all plots at the site. Perhaps use a mixture model?

Explore how to combine the total fruit equivalents with undamaged/damaged fruit counts.

```
model { 
# hyperpriors
for(i in 1:nsites){
alphaF[i] ~ dnorm(0, .001)
}

for(i in 1:nyears){
betaF[i] ~ dnorm(0, .001)
}

for(j in 1:nsites){
for(k in 1:nyears){
gammaF[j,k] ~ dnorm(0, 0.001)
rF[j,k] ~ dgamma(.001,.001)
}
}

# likelihoods
for (i in 1:N1){
lambdaF[i] = exp(alphaF[siteFec[i]] + betaF[yearFec[i]] + gammaF[siteFec[i],yearFec[i]])
y1[i] ~ dnegbin(rF[siteFec[i],yearFec[i]]/(rF[siteFec[i],yearFec[i]]+lambdaF[i]),rF[siteFec[i],yearFec[i]])

# simulated data for posterior predictive checks
y1.sim[i] ~ dnegbin(rF[siteFec[i],yearFec[i]]/(rF[siteFec[i],yearFec[i]]+lambdaF[i]),rF[siteFec[i],yearFec[i]])
}

}
```

### Seeds per fruit

Use Y<sub>9</sub> to estimate seeds per fruit.

These data are processed and analyzed in `clarkiaSeedBanks/modelBuild/fecundity/seedsPerFruitModelScripts.R`.

The JAGS script for this data is in `clarkiaSeedBanks/modelBuild/jagsScriptsFecundity/seedsAllJags.R`.

```
model { 
# hyperpriors
for(i in 1:nsites){
alphaP[i] ~ dnorm(0, .001)
}

for(i in 1:nyears){
betaP[i] ~ dnorm(0, .001)
}

for(j in 1:nsites){
for(k in 1:nyears){
gammaP[j,k] ~ dnorm(0, 0.001)
rP[j,k] ~ dgamma(.001,.001)
}
}

# likelihoods
for (i in 1:N2){
log(lambdaP[i]) = alphaP[sitePhi[i]] + betaP[yearPhi[i]] + gammaP[sitePhi[i],yearPhi[i]]
y2[i] ~ dnegbin(rP[sitePhi[i],yearPhi[i]]/(rP[sitePhi[i],yearPhi[i]]+lambdaP[i]),rP[sitePhi[i],yearPhi[i]])

# simulated data for posterior predictive checks
y2.sim[i] ~ dnegbin(rP[sitePhi[i],yearPhi[i]]/(rP[sitePhi[i],yearPhi[i]]+lambdaP[i]),rP[sitePhi[i],yearPhi[i]])
}

}
```

### Transition to seed bank

Combine estimates from belowground vital rates, seedling counts, fruits per plot, fruits per plant (from plots), and seeds per fruit to estimate this. 

Explore how to do this with/without a seed bank.

