---
output: github_document
---

### Packages

I am using the packages `tidyverse` and `readxl` (documentation:
<https://readxl.tidyverse.org/>).

Load the libraries for data processing (see
<https://github.com/r-lib/rlang/issues/669> for the overwrite message I
am suppressing).

```{r}
library(tidyverse)
library(knitr)
library(ggplot2)
```

For model building I am using:

```{r}
library(rjags)
library(tidybayes)
library(MCMCvis)
```

### Basic model

$ N(t+1) = A N(t) $

where 

$ A = s_1g_1 \sigma F \phi s_0 $

### Simulate data

```{r}
# ## Hypotheses
# 
# ## Fitness and seed banks

### COPULA

library(MASS)

probabilityOfSurvival = 0.2
meanFecundity = 5

sim <- function(rho,p.survival= probabilityOfSurvival,mu.fec=meanFecundity,reps=1000){

# a and b are hyperparameters of the binomial distribution
a = 100
b = p.survival

# alpha and beta are hyperparameters of the negative binomial distribution
alpha =  100
beta =  mu.fec

# Defines the temporal correlation between the two parameters.
rho = rho

# Generates standard multivariate normal data with correlation structure defined by rho.
Z <- mvrnorm(n=reps,mu=c(0,0), 
             matrix(data=c(1,rho,rho,1),
                    nrow=2,ncol=2))

# Apply the Normal CDF function to Z to obtain data that is uniform on the interval [0,1], but still correlated.
U = pnorm(Z)

# x is gamma distributed
X = qbinom(U[,1],size=a,prob=b) 

# y is beta distributed
Y = qnbinom(U[,2],size=alpha,mu=beta) 

X = cbind(X,Y)
return(X)
}

par(mfrow=c(3,2))
X=sim(rho=-.5)
arm::discrete.histogram(X[,1]/100,xlim=c(0,1),main="sigma");abline(v=probabilityOfSurvival,col='red',lwd=2,lty='dotted')
arm::discrete.histogram(X[,2],main="fec");abline(v=meanFecundity,col='red',lwd=2,lty='dotted')
X=sim(rho=0)
arm::discrete.histogram(X[,1]/100,xlim=c(0,1),main="sigma");abline(v=probabilityOfSurvival,col='red',lwd=2,lty='dotted')
arm::discrete.histogram(X[,2],main="fec");abline(v=meanFecundity,col='red',lwd=2,lty='dotted')
X=sim(rho=.5)
arm::discrete.histogram(X[,1]/100,xlim=c(0,1),main="sigma");abline(v=probabilityOfSurvival,col='red',lwd=2,lty='dotted')
arm::discrete.histogram(X[,2],main="fec");abline(v=meanFecundity,col='red',lwd=2,lty='dotted')

par(mfrow=c(1,3))
X=sim(rho=-.5)
plot(X[,1],X[,2],xlim=c(0,60),ylim=c(0,20),pch=16,cex=.5)
X=sim(rho=0)
plot(X[,1],X[,2],xlim=c(0,60),ylim=c(0,20),pch=16,cex=.5)
X=sim(rho=.5)
plot(X[,1],X[,2],xlim=c(0,60),ylim=c(0,20),pch=16,cex=.5)
```

Start with a set of mean values that projecs a population growth rate just above 1.

## Resample

Plot global mean and per year means

```{r}

# get global distribution
dataGlobal = sim(rho=0)
data = dataGlobal
dataGlobal = as.data.frame(dataGlobal)

# get distribution for each year by sampling mean from global distribution
years = list()

for(i in 1:5){
  years[[i]] = cbind(year=i,sim(rho=0,p.survival=data[i,][1]/100,mu.fec=data[i,][2],reps=250))
}

dataGroup = do.call(rbind.data.frame, years) %>% dplyr::mutate(year=as.factor(year))


# plots

ggplot(data=dataGlobal) +
  geom_histogram(aes(x=X/100),bins=100) +
  geom_vline(xintercept=probabilityOfSurvival,col='red') +
    theme_bw() +
  xlim(c(0,1))

ggplot() +
 #   geom_histogram(data=dataGlobal,aes(x=X/100),bins=100) +
  geom_vline(xintercept=probabilityOfSurvival,col='red') +
 geom_histogram(data=dataGroup,aes(x=X/100,fill=year),bins=100,alpha=.2,position='identity') +
    theme_bw() +
  xlim(c(0,1))


ggplot(data=dataGlobal) +
  geom_histogram(aes(x=Y),bins=100) +
  geom_vline(xintercept=meanFecundity,col='red') +
    theme_bw()

ggplot() +
  geom_vline(xintercept=meanFecundity,col='red') +
 geom_histogram(data=dataGroup,aes(x=Y,fill=year),bins=100,alpha=.2,position='identity') +
    theme_bw()

```

```{r}

# get global distribution
dataGlobal = sim(rho=0)
data = dataGlobal
dataGlobal = as.data.frame(dataGlobal)

# get distribution for each year by sampling mean from global distribution
years = list()

for(i in 1:6){
  val = sample(1:dim(dataGlobal)[1],1,replace=TRUE)
  years[[i]] = cbind(year=i,sim(rho=0,p.survival=data[val,][1]/100,mu.fec=data[val,][2]))
}

dataGroup = do.call(rbind.data.frame, years) %>% dplyr::mutate(year=as.factor(year))

ggplot() +
  geom_point(data=dataGlobal,aes(x=X/100,y=Y),size=0.5,col='red') +
  theme_bw() 

ggplot() +
  geom_point(data=dataGlobal,aes(x=X/100,y=Y),size=0.5,col='red') +
  geom_point(data=dataGroup,aes(x=X/100,y=Y),size=0.5,alpha=0.5) +
  facet_wrap(~year) +
  theme_bw() 

```

### Estimation

```{r}
set.seed(1)

# get global distribution
dataGlobal = sim(rho=0)
data = dataGlobal
dataGlobal = as.data.frame(dataGlobal)

# get distribution for each year by sampling mean from global distribution
years = list()

for(i in 1:6){
  val = sample(1:dim(dataGlobal)[1],1,replace=TRUE)
  years[[i]] = cbind(year=i,sim(rho=0,p.survival=data[val,][1]/100,mu.fec=data[val,][2]))
}

dataGroup = do.call(rbind.data.frame, years) %>% dplyr::mutate(year=as.factor(year))

names(dataGroup) = c("year","fruitplNumber","sdno")
dataGroup$seedlingNumber = 100
dataGroup$site = as.factor(1)
```
Tidy up

```{r}
dataForBayes <- tidybayes::compose_data(dataGroup)
```

 Set JAGS parameters and random seed



```{r}
# scalars that specify the 
# number of iterations in the chain for adaptation
# number of iterations for burn-in
# number of samples in the final chain

n.adapt = 3000
n.update = 5000
n.iterations = 1000
n.thin = 1

set.seed(10)
dir = c("/Users/Gregor/Dropbox/clarkiaSeedBanks/library/")

paste0(c("mu0_","sigma0_"),c("s","f"))
paramNames = c("mu0_","sigma0_")
dataNames = c("s","f")

hyperpriors <- as.vector(outer(paramNames, dataNames, paste0))
priors <- c("sigma_s","r_f")

aList <- list(0,.5,
     0,.5,
     matrix(rep(.5,1,6),nrow=1,ncol=6),
     matrix(10, nrow=1,ncol=6))
names(aList) = c(hyperpriors,priors)

# set inits for JAGS
inits = list(aList,aList,aList)

# # Call to JAGS
# 
# # tuning (n.adapt)
# jm = jags.model(file = paste0(dir,"simulationJAGS.R"), data = dataForBayes, inits = inits,
#                 n.chains = length(inits), n.adapt = n.adapt)

 jm = jags.model(file = "~/Dropbox/clarkiaSeedBanks/library/simulation/fullModelScript.txt", data = dataForBayes, inits = inits,
                 n.chains = length(inits), n.adapt = n.adapt)

```

```{r}
# burn-in (n.update)
update(jm, n.iter = n.update)

parsToMonitor_s = c("theta_s","mu0_s","sigma0_s","mu_s","sigma_s")
parsToMonitor_f = c("lambda_f","mu0_f","sigma0_f","gamma_f","r_f")
parsToMonitor_deriv = c("p0_s","p_s","p0_f","p_f")

# chain (n.iter)
samples.rjags = coda.samples(jm, 
                             variable.names = c(parsToMonitor_deriv), 
                             n.iter = n.iterations, thin = n.thin)


fileDirectory<- c("/Users/Gregor/Dropbox/dataLibrary/")
dir.create(file.path(fileDirectory), showWarnings = FALSE)
# 
saveRDS(samples.rjags,file=paste0(fileDirectory,"simulationSamples.rds"))
saveRDS(dataGroup,file=paste0(fileDirectory,"simulationData.rds"))
saveRDS(dataForBayes,file=paste0(fileDirectory,"simulationDataBayes.rds"))

MCMCsummary(samples.rjags)
```

```{r}
samples.rjags %>%
  tidybayes::recover_types() %>%
  tidybayes::spread_draws(p0_s,p0_f) %>%
  dplyr::summarise(med = median(p0_f), 
                   ci.lo = quantile(p0_f,probs=0.025), 
                   ci.hi = quantile(p0_f,probs=0.975),
                   ci.lo2 = quantile(p0_f,probs=0.25), 
                   ci.hi2 = quantile(p0_f,probs=0.75)
  )

samps = samples.rjags %>%
  tidybayes::recover_types() %>%
  tidybayes::spread_draws(p0_s,p0_f) 
 
ggplot() +
  geom_point(data=dataGlobal,aes(x=X/100,y=Y),size=0.5,col='red') +
  theme_bw() +
  geom_segment(aes(x=probabilityOfSurvival,xend=probabilityOfSurvival,y=0,yend=meanFecundity)) +
  geom_segment(aes(x=0,xend=probabilityOfSurvival,y=meanFecundity,yend=meanFecundity))

ggplot() +
  geom_point(data=samps,aes(x=p0_s,y=p0_f),size=0.5,col='red') +
  theme_bw()+ylim(c(0,250)) +
  geom_segment(aes(x=probabilityOfSurvival,xend=probabilityOfSurvival,y=0,yend=meanFecundity)) +
  geom_segment(aes(x=0,xend=probabilityOfSurvival,y=meanFecundity,yend=meanFecundity))

ggplot(data=samps) +
  geom_histogram(aes(x=p0_s),bins=100) +
  geom_vline(xintercept=probabilityOfSurvival,col='red')

ggplot(data=samps) +
  geom_histogram(aes(x=p0_f),bins=100) +
  geom_vline(xintercept=meanFecundity,col='red') +xlim(c(0,100))
```
```{r}
summaryBayes <- samples.rjags %>%
  tidybayes::recover_types(data) %>%
  tidybayes::spread_draws(p_f[site,year]) %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(med = median(p_f), 
                   ci.lo = quantile(p_f,probs=0.025), 
                   ci.hi = quantile(p_f,probs=0.975),
                   ci.lo2 = quantile(p_f,probs=0.25), 
                   ci.hi2 = quantile(p_f,probs=0.75)
  ) %>% dplyr::mutate(year = as.factor(year))
  
summaryOG <- dataGroup %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(mu=mean(sdno))

summaryBind<- summaryBayes %>% dplyr::left_join(summaryOG,by="year")

plot(summaryBind$med,summaryBind$mu,xlim=c(0,10),ylim=c(0,10));
segments(x0=summaryBind$med,y0=summaryBind$ci.lo,y1=summaryBind$ci.hi)
abline(a=0,b=1)

```

```{r}
summaryBayes <- samples.rjags %>%
  tidybayes::recover_types(data) %>%
  tidybayes::spread_draws(p_s[site,year]) %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(med = median(p_s), 
                   ci.lo = quantile(p_s,probs=0.025), 
                   ci.hi = quantile(p_s,probs=0.975),
                   ci.lo2 = quantile(p_s,probs=0.25), 
                   ci.hi2 = quantile(p_s,probs=0.75)
  ) %>% dplyr::mutate(year = as.factor(year))
  
summaryOG <- dataGroup %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(mu=mean(fruitplNumber)/100)

summaryBind<- summaryBayes %>% dplyr::left_join(summaryOG,by="year")

plot(summaryBind$med,summaryBind$mu,xlim=c(0,1),ylim=c(0,1));
segments(x0=summaryBind$med,y0=summaryBind$ci.lo,y1=summaryBind$ci.hi)
abline(a=0,b=1)

```
